---
title: "blood_analysis"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{blood_analysis}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

### Load package and data
```{r setup,warning=FALSE,message=FALSE}
library(FateMapper)
load(system.file("extdata", "blood_csc.rda", package = "FateMapper"))
```

### Data introduction
The blood lineage tracing dataset is represented as a dataframe with four columns. Each row in the dataset corresponds to a single cell, where the first and second columns contain the embedding information extracted from scRNA-seq data. The third column provides the cell-related barcode derived from lineage tracing data. Lastly, the fourth column represents the annotated cell type of each cell.
```{r}
head(cell_fate)
```

### clone profile
Firstly, we map the clone expression in pseudo-bulk cell types through fate_mapping() function
```{r}
barcode_freq = fate_mapping(cell_fate,show_row = F,cluster_cols = F,order_use = 
                              c('HSC','MPP','CMP','CLP'))
```

### cell type similarity analysis based on clone expression
Next, we calculate the spearman correlation between each cell type based on clone expression to estimate the lineage relationships between cell types.
```{r,message=FALSE}
ct_similarity = cell_type_fate_similartiy(cell_fate[,3:4])
```

### Building lineage tree
Building lineage tree based on neighbor-joining method.
```{r}
lineage_tree(cell_fate[,3:4])
```

### Clone fate bias analysis
Then, we calculate the clone fate bias based on Fisher-Exact test. Fate bias represent which fate/cell_type the clone is enriched for (e.g. Which cell type does this clone mainly develop into).
In this analysis we calculate the clone fate bias for CMP
```{r}
CMP_bias = clone_fate_bias(cell_fate[,3:4],fate_use = 'CMP')
head(CMP_bias)
```

### Clone embedding
Then we visualize the CMP bias clone on the embedding plot
```{r}
plot_clone_embedding(CMP_bias[3,1],cell_fate,colors = c(rgb(200/255,200/255,200/255),
                                             rgb(239/255,153/255,81/255)))
```
